---
title: "Computational Session 1 Day 4"
author: "Scout Leonard"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(palmerpenguins)
library(tidyverse)
```

#Write a For Loop!

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

consectutive_quarters <- vector(mode = "numeric", length = length(quarter_splits)-1)

for (i in seq_along(quarter_splits)-1) {
  t <- quarter_splits[i] + quarter_splits[i+1]
  consectutive_quarters[i] <- t
}

consectutive_quarters
```

# Basic Function Structure Practice

```{r}
birddog_sum <- function(bird, dog) {
  
  if(bird > 50){
    stop("Birds can't be greater than 50!")
  }
  
  if(bird<10){
    message("Cool little bird")
  }
  
  if(dog<0){
    warning("Can a dog be less than 0 pounds?")
  }
  
  add_birddog <- bird + dog
  return(add_birddog)
}

bird_100_dog_5 <- birddog_sum(bird = 11, dog = 5)
#storing a value

birddog_sum(bird = 25, dog = 10)

birddog_sum(bird = 4, dog = 5)

birddog_sum(bird = 6, dog = -8)


```

#### function name <- elements
#### the job to do


```{r}
say_hello <- function(friend, weekday) {
  print(paste("Hello", friend, "have a great", weekday))
}

say_hello(friend = "Tina", weekday = "Thursday")
```

### Print versus return example 

```{r}
find_multiple <- function (a,b) {
  my_multiplication <- a*b
  print(my_multiplication)
}

find_multiple(a= 10, b= 6)

find_multiple(a= 10, b=6) +12

#Usually if you are doing something like this you store the value, as below: 
my_final_value <- find_multiple(a= 10, b=12)
my_final_value
```

### Functions with Conditions 

```{r}
animal_age <- function(animal, age) {
  
  if(!animal %in% c("dog", "goat")){
    stop("Animal must be a dog or a goat.")
  }
  
  if(animal == "dog") {
    creature_age <- age * 7
  }
  
  else if(animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
}

animal_age(animal = "dog", age = 13)
# animal_age(animal = "urchin", age = 13) this would break! 

```

```{r}
adelie_gentoo <- penguins %>%
  filter(species %in% c("Adelie", "Gentoo"))

# In the filter column, do any elements match the things from our vector??
```

```{r}
adelie_gentoo_wrong <- penguins %>% 
  filter(species == c("Adelie", "Gentoo"))

# Looking for exact matches that have the order Adelie, adn then Gentoo
```

### Function with For Loops 

#### Knowing that you can wrote `colnames(of some dataframe)`, update the function so that for each iteration it returns a phrse: "The mean value of _______ is ________. 

```{r}
df_means <- function(df) {
  for (i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

df_means(df = mtcars)
```
```{r}
# Make a subset of only structural size variables from the penguins dataset for Chinstrap 
#Filter is for rows, select is for columns

chinstrap_data <- penguins %>% 
  filter(species == "Chinstrap") %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

df_means(df = chinstrap_data)

df_means(penguins)
```

### Condition to check the class of an input
```{r}
lunchtime <- function(tacos, price_per) {
  
  if (is.numeric(tacos) == FALSE) {
    stop("You must input a number of tacos.")
  }
  
  tot_price <- tacos * price_per
  return((tot_price))
}

# lunchtime(tacos = "bananas", price_per = 2.50)
lunchtime(tacos = 4, price_per = 2.50)
```

#### An example where I use values from a dataframe to calculate something based on user input 

```{r}
#Calculate the groundwater transport rate

gw_rate <- function (site) {
  
  gw_slope_depth <- data.frame(sitename = c("mountain", "prairie", "desert", "beach"), 
                               depth = c(32, 41, 63, 2), 
                               slope = c(11.2, 0.4, 0.8, 2.6))
  
  site_select <- dplyr::filter(gw_slope_depth, sitename == "site")
  
  transport_rate <- 1.4 * site_selsct$slope + 3.6 * site_seclect$depth
  
  return(transport_rate)

}

#gw_rate(site = "prairie")
```

### Now using switch

```{r}
gw_rate_switch <- function(site) {
  
  gw_slope_depth <- switch(site,
                           "mountain" = c(32, 11.2),
                           "prairie" = c(41, 0.4),
                           "desert" = c(63, 0.8),
                           "beach" = c(2, 2.6))
  
  transport_rate <- 1.4 * gw_slope_depth[2] + 3.6 * gw_slope_depth[1]
  
  return(transport_rate)
}

gw_rate_switch("beach")

```


### Logistic Growth

```{r}

logistic_growth <- function(K, N0, r, time) {
  Nt <- K / (1 + ((K - N0) / N0) * exp(-r * time))
  print(Nt)
}

logistic_growth(K = 5000, N0 = 200, r = 0.43, time = 6)

time_vec <- seq(from = 0, to = 35, by = 0.1)
bacterial_growth <- logistic_growth(K = 5000, N0 = 200, r = 0.43, time = time_vec)

bacterial_df <- data.frame(time_vec, bacterial_growth)

ggplot(data = bacterial_df, aes(x = time_vec, y = bacterial_growth)) +
  geom_point()

```


For all growth rates, we want to estimate the population at all times. 
```{r}
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
time_vec <- seq(from = 0, to = 35, by = 0.1)

out_matrix <- matrix(ncol = length(r_seq), nrow = length(time_vec))

# N0 = 100, K = 6000

for (i in seq_along(r_seq)) {
  for(j in seq_along(time_vec)) {
    pop <- logistic_growth(K = 6000, N0 = 100, r = r_seq[i], time = time_vec[j])
    out_matrix[j,i] <- pop
  }
}

out_matrix
```

#### Data Wrangling of the Out Matrix 

```{r}
out_df <- data.frame(time = time_vec, out_matrix)

# Update column names
colnames(out_df) <- c("time", paste0("gr_", r_seq))

#Pivot Longer: takes a wide format data frame and compresses it into a longer format. Long form/tidy version of the data 


out_df_long <- out_df %>% 
  pivot_longer(-time, names_to = "growth_rate", values_to = "population")

# Make in ggplot a line graph for population (y axis) over time (x axis), with color of line depending on growth rate

ggplot(data = out_df_long, aes(x = time, y = population)) + 
  geom_line(aes(color = growth_rate))
```


